---
title: "Creating Mouse Panel"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

#### So this is a script into how I made the Comprehensive Mouse panel describing each of the steps along the way.
#### Most steps are noted, but some are just described here



##### Loadign in the necessary libraries, functions and databases
```{r setup2_libraries}

source("/home/kevhu/scripts/fastReadFile.R")
library(stringr)
library(org.Hs.eg.db)
library(GenomicFeatures)
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(readxl)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)
library(biomaRt)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(AnnotationHub)
library(Biostrings)
library(GenomicRanges)
library(ensembldb)
library(EnsDb.Mmusculus.v79)
library(EnsDb.Hsapiens.v86)


ah <- AnnotationHub()
edb <- ah[["AH64944"]]

# function here is for quickly converting human gene symbols to mouse gene symbols - though it seems most mouse gene smybols are just lowercase versions of their human ones

convertHumanGeneList <- function(x){
  
  require("biomaRt")
  human = useMart("ensembl", dataset = "hsapiens_gene_ensembl")
  mouse = useMart("ensembl", dataset = "mmusculus_gene_ensembl")
  
  genesV2 = getLDS(attributes = c("hgnc_symbol"), filters = "hgnc_symbol", values = x , mart = human, attributesL = c("mgi_symbol"), martL = mouse, uniqueRows=T)
  return(genesV2)
}

```

##### So inital way to get genes to target for the mouse panel is to grab gene lists from contemporary targeted panels i.e oncomineV3 and MSK-IMPACT


```{r pressure, echo=FALSE}

ccpOncominev3 <- as.character(unlist(read.table("/home/kevhu/data/oncomineV3")))
mskGeneList <- read_xlsx("/home/kevhu/data/mskImpactGeneListSuppTab4.xlsx", skip = 2)
mskGeneList2 <- unique(mskGeneList$Gene)

mskHotspots <- read.table("/home/kevhu/data/hotspots_msk.txt",sep = "\t", header = TRUE, stringsAsFactors = FALSE)

tumorSupressors <- c("TP53","RB1", "WT1", "NF1", "NF2", "APC", "TSC1", "TSC2", "SMAD4",
                     "BRCA1","BRCA2", "PTEN", "STK11", "MSH2", "MLH1", "CDH1", 
                     "VHL", "CDKN2A", "PTCH1", "MEN1")

tumorSupressors[which(tumorSupressors %in% mskGeneList2)]
combinedGenesList <- unique(c(mskGeneList2, ccpOncominev3))
mouseCancerGenes <- convertHumanGeneList(combinedGenesList)

### maede list of conversions not done by the database utilized by the function
googledConversions <- c("H3f3a", "Cdkn2a", "Xiap", "Amer1", "P21", "Bbc3", "Kmt2d", 
                        "Kmt2a", "Mre11a", "Gnaq", "Kmt2c", "Prkn", "Notch4", "Hist1h3b",
                        "Npm1", "Abraxas1", "Cop1", "Tent5c", "Mycl")

tmp1 <- combinedGenesList[-which(combinedGenesList %in% mouseCancerGenes$HGNC.symbol)]
tmp1 <- tmp1[-which(tmp1 == "-")]

### order is a bit swapped so needed to fix it
tmp2 <- cbind(tmp1, rev(googledConversions))
colnames(tmp2) <- colnames(mouseCancerGenes)
mouseCancerGenesV2 <- rbind(mouseCancerGenes, tmp2)
mouseCancerGenesV3 <- mouseCancerGenesV2[!duplicated(mouseCancerGenesV2$HGNC.symbol),]
googleSet2 <- mouseCancerGenesV2[duplicated(mouseCancerGenesV2$HGNC.symbol),]
mouseCancerGenesV3$MGI.symbol[which(mouseCancerGenesV3$HGNC.symbol == "MYC")] <- "Myc"
mouseCancerGenesV3$MGI.symbol[which(mouseCancerGenesV3$HGNC.symbol == "SPOP")] <- "Spop"
mouseCancerGenesV3$MGI.symbol[which(mouseCancerGenesV3$HGNC.symbol == "EIF1AX")] <- "Eif1ax"
mouseCancerGenesV3$MGI.symbol[which(mouseCancerGenesV3$HGNC.symbol == "MYCN")] <- "Mycn"
mouseCancerGenesV3$MGI.symbol[which(mouseCancerGenesV3$HGNC.symbol == "PMS2")] <-"Pms2"

```

##### After getting the different gene lists from panels, I looked for cancer hotspots through COSMIC and oncotarget (MSK) - loading COSMIC table and only keeping hotspots found in both. Hashed out code are data tables are made of the mutation data. 

```{r hotspot_filtering}
cosmicTable <- faster.readfile(x = "/home/kevhu/data/CosmicMutantExport.tsv", y= 0,sep = '\t')
cosmicTable2 <- cosmicTable[-which(cosmicTable$Mutation_genome_position == ""),]
cosmicTableAAsubset <- cosmicTable2$Mutation_AA
cosmicTableAAsubset <- str_remove_all(cosmicTableAAsubset, "p\\.*")
cosmicTableAAsubset <- gsub(".{1}$", "", cosmicTableAAsubset)
tableToMatch <- paste0(cosmicTable2$Gene_name, cosmicTableAAsubset)
keysToMatch <- paste0(mskHotspots$Gene, mskHotspots$Residue)


cosmicTable3 <- cosmicTable2[match(keysToMatch, tableToMatch),]

mskHotspots$GenomicLocation <- cosmicTable3$Mutation_genome_position
mskHotspots$MutStrand <- cosmicTable3$Mutation_strand_SNP
mskHotspots$cosmicID <- cosmicTable3$Mutation_ID
mskHotspots.Cosmic <- mskHotspots[-which(is.na(cosmicTable3$Gene_name)),]

#write.table(mskHotspots.Cosmic,"/home/kevhu/data/20190201mouseHotspots.txt",
#            row.names = FALSE, col.names = TRUE, quote = FALSE, sep = "\t")
```

##### Just more steps for filtering to the final list targeted by panel
```{r hotspot_filtering2}

### adding annotations for the targets
targetAnno <- rep("Oncogene", nrow(mouseCancerGenesV3))
targetAnno[which(mouseCancerGenesV3$HGNC.symbol %in% tumorSupressors)] <- "TumorSupressorGene"
mouseCancerGenesV3$Annotation <- targetAnno
annotatedAmplicon <- rep("CopyNumber", nrow(mouseCancerGenesV3))
annotatedAmplicon[which(mouseCancerGenesV3$HGNC.symbol %in% unique(mskHotspots.Cosmic$Gene))] <- "CopyNumber/Hotspot"
annotatedAmplicon[which(toupper(mouseCancerGenesV3$MGI.symbol) %in% unique(mskHotspots.Cosmic$Gene))] <- "CopyNumber/Hotspot"
annotatedAmplicon[which(mouseCancerGenesV3$HGNC.symbol %in% tumorSupressors)] <- "AllExons"
mouseCancerGenesV3$ampliconPurpose <- annotatedAmplicon

#write.table(mouseCancerGenesV3,"/home/kevhu/data/20190129preliminaryMousePanelListCosmic.txt", sep = "\t", col.names = TRUE, row.names = FALSE,
#            quote = FALSE)

```


##### The next few steps are for deciding where in the mouse genome these hotspots lie in. I.e find conversion of these hotspots from human to mouse. I recall trying to use liftOver - from hg38 to mm10. There are chainfiles made for this purpose on UCSC. Generally these are done for larger regions and within a species, so unsurprisingly when tested these things failed. One approach I tried was to BLAST but that didn't yield too much - even tried it with polypeptide patterns surrrounding the hotspot; also was hard to figure out which of the many annnotations of the proteins to use. After a bit of reading, decided to have take canonical proteins - not sure why this wasn't obvious to me in the beggining. List taken from uniprot 

```{r}
### read in the table of canonical proteins and then reduce the list to ones of TSG since oncogenes are fully tiled in the design of the panel

uniProtCanonical <- read.table("/home/kevhu/data/uniprotCanonicalList.txt", skip = 1, header = FALSE,
                               stringsAsFactors = FALSE, sep = "\t")
tmpTable <- mouseCancerGenesV3
tmpTable$uniProt <- uniProtCanonical$V2[match(mouseCancerGenesV3$HGNC.symbol, uniProtCanonical$V1)]

tmpTable$HGNC.symbol[which(is.na(tmpTable$uniProt))]
tmpTable$uniProt[22] <- "P12524"
tmpTable$uniProt[305] <- "P46100"
tmpTable$uniProt[351] <- "Q6UWZ7"
tmpTable <- tmpTable[-348,]

tmpTable$uniProt[which(tmpTable$ampliconPurpose == "CopyNumber/Hotspot")]

matchIdxTmp <- NULL
for(i in tmpTable$uniProt[which(tmpTable$ampliconPurpose == "CopyNumber/Hotspot")]){
  print(i)
  a <- which(grepl(i, blast.res$queryID))
  matchIdxTmp <- c(matchIdxTmp, a)
}


```


##### Updating annotations of some genes
```{r}

### changing some of the labels and changing classification of oncogenes into TSGs
mouseCancerGenesV4 <- mouseCancerGenesV3
mouseCancerGenesV4$ampliconPurpose[which(mouseCancerGenesV4$ampliconPurpose == 'AllExons')] <- "CDS"
oncoToTsg <- c("Smad2", "Tgfbr2", "Arid1a", "Sox9", "Chek2", "Pms1", "Pms2", "Notch1", "Pole", "Axin1")

mouseCancerGenesV4$ampliconPurpose[which(mouseCancerGenesV4$MGI.symbol %in% oncoToTsg)] <- "CDS"
mouseCancerGenesV4$Annotation[which(mouseCancerGenesV4$MGI.symbol %in% oncoToTsg)] <- "TumorSupressorGene"

### adding other tsgs not on mouse list
tsg.hgnc <- c("DCC", "CYLD", "PTCH2")
tsg.mgi <- c("Dcc", "Cyld", "Ptch2")
tsg.purpose <- rep("CDS",3)
tsg.anno <- rep("TumorSurpressorGene", 3)
tsg.tmp <- cbind(tsg.hgnc, tsg.mgi, tsg.anno, tsg.purpose)
colnames(tsg.tmp) <- colnames(mouseCancerGenesV4)
mouseCancerGenesV4 <- rbind(mouseCancerGenesV4, tsg.tmp)

### same for oncogenes
mouseCancerGenesV4 <- rbind(mouseCancerGenesV4, c("GLI1", "Gli1", "TumorSupressorGene", "CDS"))



### now I filter out the cosmic list by only oncogenes -> helps reduce amount of things not found from blast results 
### copying code from above so I can just blast using the smaller list
mouseHotspotGenes <- mouseCancerGenesV4$HGNC.symbol[which(mouseCancerGenesV4$ampliconPurpose == "CopyNumber/Hotspot")]
mskHotspots.Cosmic.onlyOncogenes <- mskHotspots.Cosmic[which(mskHotspots.Cosmic$Gene %in% mouseHotspotGenes),]

```


##### looking at polypeptides of proteins
```{r}

#### need to check if I need to load any of these files below to get same results .... ?
uniProtCanonical <- read.table("/home/kevhu/data/20190214wholeCanonicalMouseProt.txt", skip = 1, header = FALSE,
                               stringsAsFactors = FALSE, sep = "\t")
load("/home/kevhu/data/20190214mskHotspots.onlyOnco.df.Robj")
mouseCancerGenesV4 <- read.table("/home/kevhu/data/20190214mouseCancerGenesV4.txt", header = TRUE,
                                 stringsAsFactors = FALSE, sep = "\t")

matchedVars <- mskHotspots.Cosmic.onlyOncogenes$Residue[match(uniProtCanonical$V1, mskHotspots.Cosmic.onlyOncogenes$Gene)]
uniProtCanonical$residueLocation <- matchedVars

```

